% !TeX root = ../main.tex

\chapter{Background and Related Work}\label{chapter:literature-review}

\section{Contract Lifecycle Management}\label{sec:lit:clm}

Contract Lifecycle Management (CLM) denotes the systematic management of a
contractual agreement across six canonical stages: request, author, negotiate,
execute, monitor, and renew or retire~\cite{worldcc2025benchmark}.
Each transition produces auditable artefacts — drafts, redlines, signature blocks,
obligation registers — and each stage admits distinct algorithmic interventions,
from clause assembly at the authoring stage to obligation extraction at the
monitoring stage. Figure~\ref{fig:clm-lifecycle} visualises the lifecycle;
the NLP-heavy stages are highlighted.

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[
    stage/.style={rectangle, draw=TUMBlue, fill=TUMAccentLightBlue!20,
                  rounded corners=2pt, minimum width=1.55cm, minimum height=0.55cm,
                  align=center, font=\scriptsize},
    >=stealth, thick
  ]
    \node[stage] (req)  at (0.0,0) {Request};
    \node[stage] (auth) at (2.1,0) {Author};
    \node[stage] (neg)  at (4.2,0) {Negotiate};
    \node[stage] (exec) at (6.3,0) {Execute};
    \node[stage] (mon)  at (8.4,0) {Monitor};
    \node[stage] (ren)  at (10.5,0) {\shortstack{Renew/\\Retire}};
    \draw[->] (req)  -- (auth);
    \draw[->] (auth) -- (neg);
    \draw[->] (neg)  -- (exec);
    \draw[->] (exec) -- (mon);
    \draw[->] (mon)  -- (ren);
    \draw[->, dashed]
      (ren.south) .. controls (10.5,-1.15) and (2.1,-1.15) ..
      node[below, font=\scriptsize] {amend / renew}
      (auth.south);
    % Highlight NLP-heavy stages
    \draw[TUMAccentOrange, thick, rounded corners=3pt]
      (5.25,-0.50) rectangle (9.45, 0.50);
    \node[font=\scriptsize, TUMAccentOrange] at (7.35,-0.68) {NLP sub-tasks};
  \end{tikzpicture}
  \caption{CLM lifecycle. Clause extraction, risk identification, and obligation
  tracking (highlighted) occur primarily in the Execute and Monitor stages.}
  \label{fig:clm-lifecycle}
\end{figure}

World Commerce \& Contracting estimates that organisations forfeit $\approx\!9\%$
of annual contract value through poor contract management, with detection-stage
failures (missed clauses, unobserved obligations, undetected deviations) dominating
the loss distribution~\cite{worldcc2025benchmark}. Goldman Sachs places the total
addressable population of U.S.\ legal-services tasks exposed to generative AI at
approximately 44\%, with contract review repeatedly identified as the
highest-leverage subset~\cite{goldmansachs2023ai}.

The commercial CLM market is consolidated around platforms whose positioning is
documented in successive Gartner Magic Quadrants. Sirion, Icertis, Ironclad, and
DocuSign are the four vendors consistently designated Leaders, each having rebuilt
portions of their analysis stack on LLMs since 2023. Sirion markets an Extraction
Agent and a Redline Agent under an ``agentic'' architecture. Icertis ships ICI Copilots
backed by the Microsoft Azure OpenAI Service. Ironclad embeds AI Assist for redlining
and Repository AI for post-signature search. DocuSign closed its US\$165M acquisition
of Lexion in May 2024, folding Lexion's analytics into the Intelligent Agreement
Management (IAM) platform. Adjacent vendors include Spellbook (in-Word drafting on
GPT-4 class models) and Harvey, whose contract-analysis workflow serves A\&O Shearman
and PwC~\cite{harvey2023ao,harvey2023pwc}. Across this set, public product
documentation describes single-prompt LLM calls, retrieval-augmented prompts, or
task-specific fine-tunes; no Leader has publicly disclosed a deployed multi-agent
architecture for clause-level analysis.

\section{Large Language Models}\label{sec:lit:llm-background}

An LLM is a neural language model parameterised by $\theta$ that factorises a token
sequence $\mathbf{x} = (x_1,\dots,x_n)$ autoregressively:
%
\begin{equation}\label{eq:autoreg}
  p_\theta(x_1,\dots,x_T) = \prod_{t=1}^{T} p_\theta(x_t \mid x_{<t}).
\end{equation}
%
Training minimises the negative log-likelihood
$\mathcal{L}(\theta) = -\sum_{t} \log p_\theta(x_t \mid x_{<t})$
over large text corpora following the Transformer
derivation~\cite{devlin2019bert}. Modern frontier LLMs extend Eq.~\eqref{eq:autoreg}
with three post-training layers decisive for agentic deployment: supervised
instruction tuning, preference alignment via reinforcement learning from human or AI
feedback, and structured-output fine-tuning that turns free-form generation into
validated JSON and tool calls.

Table~\ref{tab:models} summarises the frontier-model releases relevant to the
experimental envelope of this thesis.

\begin{table}[h]
  \centering
  \small
  \caption{Frontier LLM releases referenced in this thesis. Context-window figures
  denote the standard production context; beta long-context modes noted parenthetically.
  Capabilities summarised from official announcements.}
  \label{tab:models}
  \begin{tabular}{p{3.2cm}p{1.7cm}p{2.0cm}p{1.6cm}p{4.2cm}}
    \toprule
    \textbf{Model} & \textbf{Provider} & \textbf{Release} &
    \textbf{Context} & \textbf{Notes} \\
    \midrule
    GPT-4           & OpenAI    & Mar 2023 & 8K / 32K  & Multimodal preview, RLHF tuned \\
    GPT-4 Turbo     & OpenAI    & Nov 2023 & 128K      & Function calling GA, JSON mode \\
    GPT-4o          & OpenAI    & May 2024 & 128K      & Native multimodal I/O \\
    GPT-5           & OpenAI    & Aug 2025 & 400K      & Unified router over fast/thinking modes \\
    Claude 3 Opus   & Anthropic & Mar 2024 & 200K      & Tool use API \\
    Claude 3.5 Sonnet & Anthropic & Jun 2024 & 200K    & Computer-use beta, artifacts \\
    Claude Sonnet 4 & Anthropic & May 2025 & 200K      & Extended thinking; MCP code-exec tools \\
    Claude Sonnet 4.5 & Anthropic & Sep 2025 & 200K (1M beta) & Agentic/tool-orchestration focus \\
    Claude Haiku 4.5 & Anthropic & Oct 2025 & 200K     & Cost-optimised agentic tier \\
    Claude Opus 4.5 & Anthropic & Nov 2025 & 200K      & SWE-bench Verified $>$80\% \\
    Gemini 1.5 Pro  & Google    & Feb 2024 & 1M        & Mixture-of-experts, long-context \\
    Gemini 2.0 Flash & Google   & Dec 2024 & 1M        & Multimodal Live API \\
    Gemini 2.5 Pro  & Google    & Jun 2025 & 1M        & Thinking budget control, MCP support \\
    Gemini 3 Pro    & Google    & Nov 2025 & 1M        & Agentic/tool-use frontier \\
    \bottomrule
  \end{tabular}
\end{table}

A parallel layer of legal-domain products consumes these frontier models. Harvey
serves A\&O Shearman, PwC, O'Melveny \& Myers, and Macfarlanes, and was reported in
2025 at valuations rising from US\$3B to US\$11B as it expanded into Vault-based
document review and Agent Builder workflows~\cite{harvey2023ao,harvey2023pwc}.
Lexis+ AI and Thomson Reuters CoCounsel embed retrieval-augmented generation over
proprietary legal corpora. None of these systems publishes peer-reviewed evaluations
of clause-level extraction on public benchmarks. ContractEval~\cite{liu2025contracteval}
is, to our knowledge, the only large-scale academic evaluation of LLMs on contract
clause extraction conducted on a public benchmark with reproducible methodology.

\section{Agentic AI and Multi-Agent Systems}\label{sec:lit:mas-background}

An LLM-based agent is a system in which one or more LLM calls are wrapped in a
control loop that observes state, plans, invokes tools, and updates
memory~\cite{yao2023react,wei2022cot}. A Multi-Agent System (MAS) replaces a single
such loop with multiple cooperating loops, each parameterised by a role, a tool
surface, and (often) a different model. Five design axes locate any MAS in the design
space; Figure~\ref{fig:mas-design-space} maps the options and marks those ablated in
this thesis.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}[
    lbl/.style={font=\small\bfseries, text=TUMBlue,
                minimum width=2.8cm, minimum height=0.42cm, anchor=west},
    used/.style={rectangle, rounded corners=1.5pt, draw=TUMBlue,
                 fill=TUMAccentLightBlue!45, font=\footnotesize,
                 minimum height=0.42cm, inner xsep=5pt, inner ysep=2pt},
    unused/.style={rectangle, rounded corners=1.5pt, draw=gray!40,
                   fill=gray!8, font=\footnotesize, text=gray!50,
                   minimum height=0.42cm, inner xsep=5pt, inner ysep=2pt},
    node distance=0.18cm
  ]
    % Row 1 – Topology
    \node[lbl] (l1) at (0, 4.0) {Topology};
    \node[used,   right=0.15cm of l1] (t1) {Hierarchical};
    \node[used,   right=0.18cm of t1] (t2) {Single-agent};
    \node[unused, right=0.18cm of t2] (t3) {Peer};

    % Row 2 – Routing
    \node[lbl] (l2) at (0, 3.22) {Routing};
    \node[used,   right=0.15cm of l2] (r1) {None};
    \node[used,   right=0.18cm of r1] (r2) {Static};
    \node[used,   right=0.18cm of r2] (r3) {LLM-mediated};
    \node[unused, right=0.18cm of r3] (r4) {Dictionary};

    % Row 3 – Specialisation
    \node[lbl] (l3) at (0, 2.44) {Specialisation};
    \node[used,   right=0.15cm of l3] (s1) {None};
    \node[used,   right=0.18cm of s1] (s2) {Per-step};
    \node[used,   right=0.18cm of s2] (s3) {Per-domain};
    \node[unused, right=0.18cm of s3] (s4) {Per-task};

    % Row 4 – Verification
    \node[lbl] (l4) at (0, 1.66) {Verification};
    \node[used,   right=0.15cm of l4] (v1) {None};
    \node[used,   right=0.18cm of v1] (v2) {Judge agent};
    \node[unused, right=0.18cm of v2] (v3) {Reflective};
    \node[unused, right=0.18cm of v3] (v4) {Ensemble};

    % Row 5 – Protocol
    \node[lbl] (l5) at (0, 0.88) {Protocol};
    \node[used,   right=0.15cm of l5] (p1) {Function call};
    \node[unused, right=0.18cm of p1] (p2) {MCP};
    \node[unused, right=0.18cm of p2] (p3) {A2A};

    % Bounding rules
    \draw[gray!20, thin] (-0.1, 0.50) -- (9.6, 0.50);
    \draw[gray!20, thin] (-0.1, 4.38) -- (9.6, 4.38);
    \draw[gray!20, thin] (2.78, 0.50) -- (2.78, 4.38);

    % Legend
    \node[used,   font=\scriptsize] at (4.2, 0.18) {\strut Ablated in this thesis};
    \node[unused, font=\scriptsize] at (7.0, 0.18) {\strut Not ablated};
  \end{tikzpicture}
  \caption{Multi-agent design space: five structural axes and their principal options.
  Shaded boxes mark options covered by at least one experimental configuration
  (B1, B4, M0, M1, M6) in this thesis. Chapter~3 specifies which configuration
  selects which option.}
  \label{fig:mas-design-space}
\end{figure}

The empirical literature has converged on five design axes along which any MAS can be
located. \emph{Topology} distinguishes hierarchical configurations (orchestrator
dispatches to subordinate workers) from peer configurations (agents communicate
without a central coordinator), as in MetaGPT's role-pipeline~\cite{hong2024metagpt}
and CAMEL's role-playing pairs~\cite{li2023camel}. \emph{Routing} distinguishes
static pipelines (fixed step order), LLM-mediated routing (a router agent selects a
downstream worker), and dictionary-based routing (a deterministic feature-to-worker
map). \emph{Specialisation granularity} ranges from per-task agents (one per CUAD
category) through per-domain agents (one for risk, IP, temporal) to per-step agents
(extract, verify, format). \emph{Verification structure} ranges from none, through
reflective self-critique~\cite{shinn2023reflexion}, multi-agent debate, judge-agent
verification, tree search~\cite{yao2023tot}, majority-vote ensembles such as
Mixture-of-Agents~\cite{wang2024moa}, and self-consistency~\cite{wang2023selfconsistency}.
The \emph{communication protocol} is the lowest layer: function calling via vendor
APIs, the Model Context Protocol (MCP) released by Anthropic in November
2024~\cite{anthropicmcp2024}, the Agent2Agent (A2A) protocol released by Google in
April 2025 and donated to the Linux Foundation in June 2025~\cite{googlea2a2025}, or
framework-specific structured messages.

Production-relevant orchestrators include AutoGen~\cite{wu2023autogen},
MetaGPT~\cite{hong2024metagpt}, LangGraph~\cite{langgraph2024}, CrewAI, and
OpenAI's Agents SDK. The five axes — topology, routing, specialisation granularity,
verification structure, and communication protocol — define the design space whose
CLM-relevant slice this thesis ablates in Chapter~3.

\section{The CUAD Benchmark}\label{sec:lit:cuad}\label{sec:cuad}

The Contract Understanding Atticus Dataset (CUAD) is the canonical public benchmark
for clause-level contract review~\cite{hendrycks2021cuad}. It consists of 510
commercial contracts annotated against 41 categories of legal clauses. The annotation
pipeline employed law students who completed 70--100 hours of contract-review training
under the supervision of practising attorneys~\cite{hendrycks2021cuad}, a provenance
restated by ContractEval as a strength relative to crowd-sourced legal
datasets~\cite{liu2025contracteval}. Table~\ref{tab:cuad-stats} reproduces the
dataset statistics.

\begin{table}[h]
  \centering
  \small
  \caption{CUAD dataset statistics~\cite{hendrycks2021cuad}.}
  \label{tab:cuad-stats}
  \begin{tabular}{ll}
    \toprule
    \textbf{Property} & \textbf{Value} \\
    \midrule
    Contracts            & 510 \\
    Clause categories    & 41 \\
    Expert annotations   & $>$13{,}000 spans \\
    Source               & EDGAR commercial filings \\
    Train / test split   & 399 / 102 \\
    Task formulation     & Span extraction + binary presence \\
    License              & CC BY 4.0 \\
    Long contracts ($>$100k chars) & 81 (used in this thesis) \\
    \bottomrule
  \end{tabular}
\end{table}

The instance count is highly imbalanced: at the (contract, category) level, the
natural prevalence of positive instances is $\approx\!30\%$, a structural property
of contract review rather than a sampling artefact. This thesis evaluates on the
long-contract subset comprising 81 contracts whose raw character length exceeds
100{,}000 characters. Two considerations motivate this choice: first, the long-contract
regime is where context-window pressure on a single LLM call is highest, so any
putative benefit of architectural decomposition should be most visible; second,
concentrating on long contracts produces an architecturally more demanding evaluation
envelope than the full CUAD distribution.
Figure~\ref{fig:cuad-tiers} illustrates the F$_1$ gradient across tiers that
motivates the domain-specialisation hypothesis (H2).

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    \begin{axis}[
      xbar,
      width=0.84\textwidth,
      height=8.0cm,
      xmin=0, xmax=1.0,
      xlabel={Approximate $F_1$ (GPT-4.1, ContractEval~\cite{liu2025contracteval})},
      xlabel style={font=\small},
      xtick={0, 0.25, 0.5, 0.75, 1.0},
      xticklabels={0, .25, .50, .75, 1.0},
      xticklabel style={font=\scriptsize},
      ytick={0,1,...,12},
      yticklabels={
        Vol.\ Restriction,      % 0  rare
        Joint IP Ownership,     % 1  rare
        Notice Period,          % 2  rare
        Uncapped Liability,     % 3  rare
        Renewal Term,           % 4  moderate
        Change of Control,      % 5  moderate
        License Grant,          % 6  moderate
        Anti-Assignment,        % 7  moderate
        Termination Conv.,      % 8  moderate
        Expiration Date,        % 9  common
        Effective Date,         % 10 common
        Parties,                % 11 common
        Governing Law           % 12 common
      },
      yticklabel style={font=\scriptsize, text width=3.0cm, align=right},
      bar width=0.32cm,
      enlarge y limits=0.05,
      legend style={
        at={(0.5,-0.16)}, anchor=north,
        font=\scriptsize, draw=gray!30, fill=white,
        legend columns=3, column sep=0.5cm
      },
    ]
      % Common tier
      \addplot[fill=TUMBlue!65, draw=TUMBlue, bar shift=0pt] coordinates {
        (0.85, 12) (0.82, 11) (0.76, 10) (0.72, 9)
      };
      % Moderate tier
      \addplot[fill=TUMAccentGray, draw=TUMDarkGray!50, bar shift=0pt] coordinates {
        (0.58, 8) (0.52, 7) (0.47, 6) (0.41, 5) (0.38, 4)
      };
      % Rare tier
      \addplot[fill=TUMAccentOrange!75, draw=TUMAccentOrange, bar shift=0pt] coordinates {
        (0.04, 3) (0.05, 2) (0.03, 1) (0.01, 0)
      };
      \legend{Common ($F_1 > 0.70$), Moderate ($0.30$--$0.70$), Rare ($< 0.15$)}
    \end{axis}
  \end{tikzpicture}
  \caption{Approximate $F_1$ scores for 13 representative CUAD categories under
  GPT-4.1 (ContractEval~\cite{liu2025contracteval}). Values are indicative; exact
  scores vary by model. Rare-tier categories achieve near-zero $F_1$ under the
  strongest single-agent baseline, directly motivating the per-domain specialist
  design evaluated in Chapter~3.}
  \label{fig:cuad-tiers}
\end{figure}

\section{Contract Analysis Methods}\label{sec:lit:methods}

The methodological lineage of automated contract analysis runs from rule-based
extractors through statistical sequence labellers to fine-tuned and prompted
Transformers. Early systems used hand-written regular expressions and gazetteers over
POS-tagged text. Linear-chain Conditional Random Fields~\cite{lafferty2001crf} and
BiLSTM--CRF taggers~\cite{lample2016neural} formalised the task as token-level
sequence labelling and dominated the LEDGAR~\cite{tuggener2020ledgar} and
ContractNLI~\cite{koreeda2021contractnli} eras. The release of BERT~\cite{devlin2019bert}
and legal-domain LEGAL-BERT~\cite{chalkidis2020legalbert} shifted the regime from
feature engineering to pre-train-and-fine-tune. CUAD's reference numbers were obtained
with RoBERTa-style extractive question answering~\cite{hendrycks2021cuad}. MAUD~\cite{wang2023maud}
extended the paradigm to merger-agreement understanding.

The LLM-prompting regime is characterised by two reproducible studies.
Better Call GPT~\cite{martin2024bettercallgpt} reported that GPT-4 with structured
prompting can match or exceed human associates on contract-review subtasks under
controlled conditions. ContractEval~\cite{liu2025contracteval} evaluated 19 LLMs
(4 proprietary, 15 open-source) on CUAD, found a persistent gap between proprietary
and open-source models, and identified the named failure mode of \emph{laziness} —
the model declines to extract any span when the clause is present. ContractEval also
documented a tension between reasoning mode and correctness: chain-of-thought-style
decoding improved Jaccard span overlap on positive instances but degraded overall
correctness, particularly on negative instances.

Table~\ref{tab:datasets} compares the main contract-analysis benchmarks.

\begin{table}[h]
  \centering
  \small
  \caption{Main contract-analysis benchmarks.}
  \label{tab:datasets}
  \begin{tabular}{lllll}
    \toprule
    \textbf{Dataset} & \textbf{Task} & \textbf{Size} & \textbf{Categories} &
    \textbf{Used here} \\
    \midrule
    CUAD~\cite{hendrycks2021cuad}           & Span extraction & 510 contracts  & 41  & \checkmark \\
    ContractNLI~\cite{koreeda2021contractnli} & Doc-level NLI  & 607 NDAs      & 17  & --- \\
    MAUD~\cite{wang2023maud}                & Reading comp.  & 152 M\&A agmt. & 152 & --- \\
    LEDGAR~\cite{tuggener2020ledgar}        & Multi-label cls.& 60k contracts  & 100+& --- \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Why multi-agent for contract analysis.}
Four properties of the CUAD-style problem motivate ablating multi-agent architectures
on this task. First, contracts in the long-context subset routinely exceed 100{,}000
characters; explicit decomposition makes per-call context tractable where single-prompt
formulations must truncate or chunk silently. Second, the 41 CUAD categories partition
into risk-oriented, IP/commercial, temporal, and consideration-oriented clusters; a
specialisation hypothesis predicts that role-conditioned agents may outperform a single
generalist on within-cluster categories. Third, legal review is recall-asymmetric: a
missed clause exposes the client to undisclosed obligations whereas a spurious extraction
is corrected during attorney review at lower cost — motivating $F_2$-weighted evaluation
(Section~\ref{sec:lit:eval}) and architectures that boost recall at controlled precision
cost. Fourth, regulated environments — the EU AI Act treats certain legal AI uses as
high-risk~\cite{euaiact2024} and GDPR requires data-protection accountability~\cite{gdpr2016}
— demand traceable per-decision provenance; multi-agent architectures with explicit
verifier or judge agents produce that trace at the architectural level.

\section{Multi-Agent Legal AI}\label{sec:lit:mas}

Table~\ref{tab:mas} codes the published multi-agent legal systems most directly
comparable to this thesis along the five design axes introduced in
Section~\ref{sec:lit:mas-background}.

\begin{table}[h]
  \centering
  \small
  \setlength{\tabcolsep}{5pt}
  \renewcommand{\arraystretch}{1.45}
  \caption{Multi-agent legal AI systems coded against the five design-space axes of
  Section~\ref{sec:lit:mas-background}. All prior systems evaluate a single architecture
  without ablation.}
  \label{tab:mas}
  \newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
  \begin{tabular}{L{2.6cm}L{2.4cm}L{2.1cm}L{2.1cm}L{2.9cm}L{2.4cm}}
    \toprule
    \textbf{System} & \textbf{Task} & \textbf{Topology} & \textbf{Routing} &
    \textbf{Specialisation} & \textbf{Verification} \\
    \midrule
    ChatLaw v2 \cite{cui2024chatlaw}
      & Legal QA (CN) & Hierarchical & LLM-mediated & Per-domain & Self-check \\
    PAKTON \cite{raptopoulos2025pakton}
      & Contract review & Hierarchical & Static & Per-step (3) & None \\
    L-MARS \cite{wang2025lmars}
      & Statutory research & Hierarchical & LLM-mediated & Per-step + retrieval & Judge agent \\
    Lawma \cite{dominguezolmedo2024lawma}
      & Legal annotation & Single-agent & N/A & Per-task (fine-tune) & None \\
    Layered CoT \cite{sanwal2025layered}
      & Statutory reasoning & Single-agent & Static & Per-step CoT layers & Self-consistency \\
    \midrule
    \textbf{This thesis}
      & CUAD long-contract & Both & All three & All three & Reflexion + judge \\
    \bottomrule
  \end{tabular}
\end{table}

ChatLaw v2~\cite{cui2024chatlaw} couples a Mixture-of-Experts router with
role-conditioned domain agents for Chinese legal consultation; verification is a
self-check pass and there is no public ablation against single-agent variants.
PAKTON~\cite{raptopoulos2025pakton} pipelines a reading agent, an analysis agent,
and a drafting agent over open-ended contract-review prompts; the system is evaluated
qualitatively and not on CUAD. L-MARS~\cite{wang2025lmars} combines a planner, a
retriever, and a judge agent on statutory research; performance is reported at system
level without architectural decomposition. Lawma~\cite{dominguezolmedo2024lawma},
while not multi-agent, provides the strongest counter-evidence to architectural
complexity in the legal domain: a fine-tuned 8B Llama 3 model exceeds GPT-4 by 20.0
accuracy points on legal annotation, and the smallest variant (135M) is competitive
with Claude 3.7 Sonnet. The layered chain-of-thought system of
Sanwal~\cite{sanwal2025layered} stays within a single agent but stacks specialist
prompts and self-consistency, reporting gains over flat prompting on statutory reasoning.

Across this set, every prior system reports performance for a single chosen architecture
without ablating it against simpler or alternative configurations on the same evaluation
envelope. The question of whether the architecture or the model is responsible for
observed performance therefore remains open. This thesis addresses it directly through
cross-architecture ablation on a fixed CUAD evaluation harness.

\section{Conditions for Multi-Agent Effectiveness}\label{sec:lit:when}

Four recent studies bound the conditions under which multi-agent systems produce gains
over single-agent baselines. Gao et al.~\cite{gao2025single} show on a battery of
reasoning and tool-use tasks that single, well-prompted LLMs frequently match or exceed
multi-agent configurations once the latter are corrected for cost and prompting parity,
and that observed multi-agent gains often disappear under stronger single-agent baselines.
Cemri et al.~\cite{cemri2025why} taxonomise multi-agent failure modes —
specification drift between agents, inter-agent verification gaps, lossy summarisation
across hand-offs — and document that these failures account for the majority of
regressions in the systems they audit. Bogavelli et al.~\cite{bogavelli2025agentarch}
run a large architectural sweep across orchestration patterns and base models and find
substantial model-family $\times$ architecture interactions: an architecture that is
optimal under one model family is often dominated under another.
Sanwal~\cite{sanwal2025layered} demonstrates that layered single-agent chain-of-thought
can recover much of the benefit attributed to multi-agent decomposition without paying
the coordination cost.

Synthesised against the working hypotheses of this thesis: (H1) a generic claim of
multi-agent superiority on CUAD long-contract extraction is unlikely to survive a
properly normalised single-agent baseline~\cite{gao2025single}, so the hypothesis is
stated and evaluated as conditional; (H2) the relative ordering of architectures will
depend on the underlying model family~\cite{bogavelli2025agentarch}, encoded as an
explicit interaction term in the experimental design; and (H3) classification-gate
cascades — a router agent that decides whether to extract before invoking an extractor
— are a natural counter to the verification-gap and laziness failures
catalogued in~\cite{cemri2025why}, motivating the M0 ablation in Chapter~3.

\section{Evaluation Methodology}\label{sec:lit:eval}\label{sec:eval}

This section consolidates the evaluation primitives used throughout the thesis.
Chapter~3 references these definitions without restating them.

\paragraph{Span and instance metrics.}
For a category $c$ on a contract $d$, let $S^*_{c,d}$ be the set of gold spans and
$\hat{S}_{c,d}$ the set of predicted spans. A predicted span $\hat{s}$ matches a gold
span $s$ when their character-level Jaccard overlap exceeds threshold $\tau{=}0.5$
following~\cite{hendrycks2021cuad}:
%
\begin{equation}\label{eq:jaccard}
  J(\hat{s}, s) = \frac{|\hat{s} \cap s|}{|\hat{s} \cup s|}.
\end{equation}
%
At the (contract, category) level, the prediction is true-positive if any predicted
span matches any gold span at threshold $\tau$, false-negative if gold spans exist
with no match, false-positive if predicted spans exist with no match, and true-negative
if both sets are empty. Precision $P$ and recall $R$ aggregate over the natural class
distribution.

\paragraph{$F_\beta$.}
Recall is privileged over precision in legal review by a factor commonly fixed at
$\beta{=}2$:
%
\begin{equation}\label{eq:fbeta}
  F_\beta = \frac{(1+\beta^2)\,P\,R}{\beta^2 P + R}, \qquad \beta=2.
\end{equation}
%
$F_1$ is reported alongside $F_2$ for compatibility with prior CUAD work.
Instance-level $F_\beta$ at the (contract, category) granularity is the primary
metric reported in Chapter~4, as it aligns with the deployment unit (a reviewer's
decision per category per contract).

\paragraph{Failure-mode metrics.}
ContractEval~\cite{liu2025contracteval} introduced \emph{laziness} as the per-model
rate at which an LLM declines to extract on instances where a gold span exists:
%
\begin{equation}\label{eq:laziness}
  \lambda = \frac{|\{(c,d): |\hat{S}_{c,d}|=0 \;\wedge\; |S^*_{c,d}|>0\}|}
                 {|\{(c,d): |S^*_{c,d}|>0\}|}.
\end{equation}
%
The dual failure mode is \emph{over-extraction} — the per-model rate of producing
spans on negative instances:
%
\begin{equation}\label{eq:overextract}
  \omega = \frac{|\{(c,d): |\hat{S}_{c,d}|>0 \;\wedge\; |S^*_{c,d}|=0\}|}
                {|\{(c,d): |S^*_{c,d}|=0\}|}.
\end{equation}
%
Equations~\eqref{eq:laziness} and~\eqref{eq:overextract} are a coupled pair:
evaluations that resample to a balanced class distribution suppress
Eq.~\eqref{eq:overextract} and hide the precision cost of recall-boosting
interventions. We therefore evaluate exclusively on the natural CUAD class distribution
and report both metrics.

\paragraph{Statistical testing.}
We pair every architecture comparison on the same contract set and report McNemar's
test for instance-level binary outcomes~\cite{mcnemar1947note} and the paired bootstrap
with $B{=}10{,}000$ resamples for span-level continuous
metrics~\cite{efron1979bootstrap,riezler2005pitfalls}. Multiple-comparison control
across the architecture grid uses Benjamini--Hochberg false-discovery-rate correction
at $q{=}0.05$~\cite{benjamini1995controlling}. Sample-size and power considerations
follow Dror et al.~\cite{dror2018hitchhikers} and Card et al.~\cite{card2020little};
the 81-contract subset gives $\geq 0.8$ power to detect absolute $F_2$ differences of
$\approx\!2.5$ percentage points at the instance level under the bootstrap.
Implementations use SciPy~\cite{virtanen2020scipy} and NumPy~\cite{harris2020array}.

\paragraph{Sampling design.}
CUAD's natural clause-presence rate is $\approx\!30\%$ positive at the instance level.
Resampling to balance positive and negative instances inflates apparent precision and
suppresses Eq.~\eqref{eq:overextract}, detaching the evaluation from any deployable
distribution where 41 categories are scanned per contract and most are absent. We
retain natural prevalence throughout~\cite{liu2025contracteval}.

\paragraph{ContractEval as direct empirical predecessor.}
ContractEval~\cite{liu2025contracteval} evaluates 19 LLMs (4 proprietary, 15
open-source) on 4{,}128 (contract, category) samples across the full 41-category CUAD
schema, holding prompt and decoding constants across models, and reporting Jaccard span
overlap, instance-level correctness, and laziness. Its principal findings are: a
persistent proprietary--open-source gap on legal extraction; a non-monotonic effect of
explicit reasoning (chain-of-thought improves Jaccard on positive instances while
degrading correctness through inflated false positives); and laziness as a robust,
model-dependent and category-dependent failure mode. Three properties of the
ContractEval design define the gaps the present thesis closes: the architectural axis
is held fixed at single-prompt single-model; model-family interactions are not
analysed; and over-extraction (Eq.~\eqref{eq:overextract}) is not tracked as a
first-class metric. We retain ContractEval's evaluation protocol —
natural class distribution, instance-level scoring, laziness reporting,
$\tau{=}0.5$ Jaccard threshold — and extend it along the architectural axis and the
dual failure-mode axis.

\paragraph{Prompt sensitivity as a validity threat.}
POSIX~\cite{chatterjee2024posix} quantifies the sensitivity of LLM outputs to
semantically equivalent prompt rewrites, reporting non-trivial sensitivity even on
well-defined tasks. Prompt sensitivity is therefore a first-order validity threat for
any architecture comparison in which architectures differ in the number, role, or
wording of agent prompts. Chapter~3 controls this threat by holding system-prompt
scaffolds fixed across architectures where the role permits, and reports a
POSIX-style sensitivity check on the architectural ablations. Observability and trace
inspection use Langfuse~\cite{langfuse2024}; orchestration uses LangGraph~\cite{langgraph2024}.

\section{Research Gap}\label{sec:lit:synthesis}

Four gaps in the published record jointly motivate this thesis. (i) No published study
performs a controlled cross-architecture ablation on CUAD-class clause extraction; the
closest predecessor~\cite{liu2025contracteval} varies model identity only.
(ii) No published study reports a model-family $\times$ architecture interaction
analysis on this task, despite the prediction in~\cite{bogavelli2025agentarch} that
such interactions dominate single-axis effects. (iii) No published study tracks both
Eq.~\eqref{eq:laziness} and Eq.~\eqref{eq:overextract} under the natural class
distribution, leaving the recall--precision asymmetry intrinsic to legal review
unobserved. (iv) Production CLM platforms — Sirion, Icertis, Ironclad, DocuSign —
disclose single-prompt or retrieval-augmented LLM deployments with no public
multi-agent ablation, leaving the architectural design space untested in the regime
where the question is commercially decisive.

We design and evaluate a slice of the multi-agent design space (topology $\times$
routing $\times$ specialisation $\times$ verification) on the long-contract CUAD
subset under the ContractEval protocol extended with Eq.~\eqref{eq:overextract} and a
model-family interaction term, in order to determine — with statistical control across
the architecture grid~\cite{benjamini1995controlling,dror2018hitchhikers} — under
which conditions multi-agent decomposition is justified for CLM deployment and under
which conditions it is dominated by a stronger single-agent baseline.
