% !TeX root = ../main.tex

\chapter{Methodology}\label{chapter:methodology}

Each architectural comparison holds one axis constant while varying one: B1 vs.\ B0
isolates chain-of-thought reasoning, M0 vs.\ B0 isolates the classification gate,
and M1 vs.\ M2 isolates routing from prompt content. This controlled design allows
performance differences to be attributed to architecture rather than model or data
variation across $5 \times 11 = 55$ runs on the 102-contract CUAD test split.

\section{Research Design}\label{sec:meth:design}

Five configurations are evaluated against 11 frontier LLMs on the 102 contracts of
the CUAD test split (character lengths 645\,--\,300{,}768; median 25{,}657),
yielding $5 \times 11 = 55$ official runs and 163{,}644
analysed predictions. Differences between configurations are minimised to a single
architectural axis per comparison (Table~\ref{tab:configs}), following the controlled
comparative protocol of \cite{dror2018hitchhikers,card2020little}. Every statistical
test is paired on the intersection of sample IDs, so performance differences are
attributable to architecture or model, not dataset drift.

\begin{table}[h]
  \centering
  \small
  \caption{Five evaluated configurations. B1 vs.\ B0 isolates chain-of-thought;
  M0 vs.\ B0 isolates the classification gate; M1 vs.\ M2 isolates routing and
  specialist decomposition (architecture ablation).
  ${}^\dagger$Artifact IDs used in the experiment repository: B0\,=\,B1,
  B1\,=\,B4, M0\,=\,M0, M1\,=\,M1, M2\,=\,M6.}
  \label{tab:configs}
  \begin{tabular}{lllll}
    \toprule
    \textbf{ID} & \textbf{Name} & \textbf{Stages} & \textbf{Routing} &
    \textbf{Specialisation} \\
    \midrule
    B0 & \cfgZS                & 1 & None        & Generalist \\
    B1 & \cfgCoT               & 1 & None        & Generalist + chain-of-thought \\
    M0 & \cfgCX                & 2 & Static gate & Generalist \\
    M1 & \cfgMA                & 2 & LLM router  & Three domain specialists \\
    M2 & \cfgCP                & 1 & None        & Union of M1 specialist prompts \\
    \bottomrule
  \end{tabular}
\end{table}

Figure~\ref{fig:pipelines} shows the message-flow diagram for each configuration.

\begin{figure}[t]
  \centering
  %
  % ---- shared TikZ styles ----
  \tikzset{
    nd/.style={rectangle, draw=TUMBlue, fill=TUMAccentLightBlue!20,
               rounded corners=1pt, minimum width=1.45cm, minimum height=0.42cm,
               align=center, font=\scriptsize},
    sp/.style={rectangle, draw=TUMAccentOrange, fill=TUMAccentOrange!15,
               rounded corners=1pt, minimum width=1.3cm, minimum height=0.38cm,
               align=center, font=\scriptsize},
    dc/.style={diamond, draw=TUMBlue, fill=TUMAccentLightBlue!20,
               aspect=2.2, align=center, font=\scriptsize,
               minimum width=1.2cm, minimum height=0.45cm},
    >=stealth, thick
  }

  % ---- Row 1: Baselines (B0, B1) ----
  \begin{subfigure}[b]{0.28\textwidth}
    \centering
    \begin{tikzpicture}[node distance=0.42cm]
      \node[nd] (i)   {Input};
      \node[nd, below=of i] (l) {LLM};
      \node[nd, below=of l] (o) {Answer};
      \draw[->] (i) -- (l);
      \draw[->] (l) -- (o);
    \end{tikzpicture}
    \caption{\cfgZS\ (B0)}
    \label{fig:arch-b1}
  \end{subfigure}
  \hfil
  \begin{subfigure}[b]{0.28\textwidth}
    \centering
    \begin{tikzpicture}[node distance=0.42cm]
      \node[nd] (i)   {Input};
      \node[nd, below=of i] (l) {LLM};
      \node[nd, below=of l] (c) {CoT};
      \node[nd, below=of c] (o) {Answer};
      \draw[->] (i) -- (l);
      \draw[->] (l) -- (c);
      \draw[->] (c) -- (o);
    \end{tikzpicture}
    \caption{\cfgCoT\ (B1)}
    \label{fig:arch-b4}
  \end{subfigure}

  \par\bigskip

  % ---- Row 2: Multi-agent (M0, M1, M2) ----
  \begin{subfigure}[b]{0.30\textwidth}
    \centering
    \begin{tikzpicture}[node distance=0.42cm]
      \node[nd] (i)  {Input};
      \node[nd, below=of i] (cl) {Classifier};
      \node[dc, below=of cl] (d) {Present?};
      \node[nd, below left=0.5cm and 0.1cm of d] (ex) {Extractor};
      \node[nd, below right=0.5cm and 0.1cm of d] (nu) {$\emptyset$};
      \node[nd, below=1.35cm of d] (o) {Answer};
      \draw[->] (i)  -- (cl);
      \draw[->] (cl) -- (d);
      \draw[->] (d.west) -| node[above left, font=\scriptsize]{yes} (ex.north);
      \draw[->] (d.east) -| node[above right, font=\scriptsize]{no}  (nu.north);
      \draw[->] (ex.south) |- (o.west);
      \draw[->] (nu.south) |- (o.east);
    \end{tikzpicture}
    \caption{\cfgCX\ (M0)}
    \label{fig:arch-m0}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.34\textwidth}
    \centering
    \begin{tikzpicture}[node distance=0.42cm]
      \node[nd] (i)  {Input};
      \node[nd, below=of i] (or) {Orchestrator};
      \node[nd, below=of or] (ro) {Router};
      \node[sp, below left=0.5cm and 0.15cm of ro]  (s1) {Risk};
      \node[sp, below=0.5cm of ro]                   (s2) {Temporal};
      \node[sp, below right=0.5cm and 0.15cm of ro] (s3) {IP/Comm.};
      \node[nd, below=1.4cm of ro] (o) {Answer};
      \draw[->] (i)  -- (or);
      \draw[->] (or) -- (ro);
      \draw[->] (ro.west)  -| (s1.north);
      \draw[->] (ro)       -- (s2.north);
      \draw[->] (ro.east)  -| (s3.north);
      \draw[->] (s1.south) |- (o.west);
      \draw[->] (s2)        -- (o);
      \draw[->] (s3.south) |- (o.east);
    \end{tikzpicture}
    \caption{\cfgMA\ (M1)}
    \label{fig:arch-m1}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.26\textwidth}
    \centering
    \begin{tikzpicture}[node distance=0.42cm]
      \node[nd] (i)  {Input};
      \node[nd, below=of i] (l) {\shortstack{LLM\\(union)}};
      \node[nd, below=of l] (o) {Answer};
      \draw[->] (i) -- (l);
      \draw[->] (l) -- (o);
    \end{tikzpicture}
    \caption{\cfgCP\ (M2)}
    \label{fig:arch-m6}
  \end{subfigure}

  \caption{Message-flow diagrams for the five configurations. Input = contract text
  + clause-category indicator. Blue nodes: generalist LLM invocations. Orange nodes:
  domain-specialist prompts. $\emptyset$ = silent no-clause return.}
  \label{fig:pipelines}
\end{figure}

\section{Dataset and Sampling}\label{sec:meth:data}

CUAD~\cite{hendrycks2021cuad} is the evaluation corpus (see Table~\ref{tab:cuad-stats}
for statistics). The full 102-contract CUAD test split is used; contract lengths
range from 645 to 300{,}768 characters (median 25{,}657), spanning short NDAs to
long master agreements. The 102-contract corpus yields
$102 \times 41 = 4{,}182$ raw (contract, category) pairs per evaluation pass, of
which $\approx\!2{,}999$ enter each run after the sampler described below.

\paragraph{Sampling strategy.}\label{sec:meth:sampling}
A \texttt{neg\_ratio=0.7} sampler retains every positive (contract, category)
pair and approximately $60\%$ of the negative pairs, yielding $\approx\!2{,}999$
samples per run with a measured positive rate of $41.5\%$. The sampler upsamples
the positive class relative to CUAD's natural $\approx\!30\%$ prior to preserve
statistical power on rare-tier categories, where positive instances are sparse;
without upsampling, fewer than ten positive instances per rare category would
remain. Aggressive resampling toward a balanced class distribution is rejected
because it suppresses the false-positive denominator and masks over-extraction
(Eq.~\ref{eq:overextract}); the chosen $41.5\%$ rate keeps both laziness
(Eq.~\ref{eq:laziness}) and over-extraction observable. Sampling is seeded
(\texttt{seed=42}) so the same (contract, category) pairs are evaluated across
all configurations, making every statistical test paired at the instance level.
Each run yields $\approx\!2{,}999$ predictions; 1{,}297 of 164{,}941 total
($0.79\%$) were excluded for malformed classification tokens.

\paragraph{Prevalence tiers.}
The 41 CUAD categories are partitioned into three tiers (\emph{common},
\emph{moderate}, \emph{rare}) by natural positive rate in the 102-contract test
split. Tier cutpoints are fixed before any model is run.

\section{Model Selection}\label{sec:meth:models}

Table~\ref{tab:models} lists the 11 frontier proprietary models in the panel. Three
families span the providers used in commercial CLM (Anthropic, Google, OpenAI), so
model-family effects predicted by~\cite{bogavelli2025agentarch} can be observed; within
each family the panel spans a compute-tier range from light (Haiku, Flash, Nano) to
heavy (Opus, Pro), enabling capability $\times$ architecture interactions to be
observed. Model identifiers are pinned to specific snapshots (see §\ref{sec:sys:stack})
for reproducibility.

\begin{table}[h]
  \centering
  \small
  \caption{Eleven frontier proprietary models evaluated. Three families; compute-tier
  range covers light-to-heavy within each family. Snapshot IDs are the exact
  provider model strings used in all API calls.}
  \label{tab:models}
  \begin{tabular}{llll}
    \toprule
    \textbf{Family} & \textbf{Model} & \textbf{Tier} & \textbf{Snapshot ID} \\
    \midrule
    \multirow{4}{*}{Anthropic}
      & Claude Haiku 4.5  & Light & \texttt{claude-haiku-4-5-20251001}  \\
      & Claude Sonnet 4   & Mid   & \texttt{claude-sonnet-4-20250514}   \\
      & Claude Sonnet 4.6 & Mid   & \texttt{claude-sonnet-4-6-20251001} \\
      & Claude Opus 4.6   & Heavy & \texttt{claude-opus-4-6}            \\
    \midrule
    \multirow{4}{*}{Google}
      & Gemini 2.5 Flash  & Light & \texttt{gemini-2.5-flash}           \\
      & Gemini 2.5 Pro    & Heavy & \texttt{gemini-2.5-pro}             \\
      & Gemini 3 Flash    & Light & \texttt{gemini-3-flash-preview}      \\
      & Gemini 3.1 Pro    & Heavy & \texttt{gemini-3.1-pro-preview}      \\
    \midrule
    \multirow{3}{*}{OpenAI}
      & GPT-4.1           & Mid   & \texttt{gpt-4.1}                    \\
      & GPT-4.1 Mini      & Light & \texttt{gpt-4.1-mini}               \\
      & GPT-4.1 Nano      & Light & \texttt{gpt-4.1-nano}               \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Inference parameters.}
All model calls used \texttt{temperature=0} for deterministic output and
\texttt{max\_tokens=4096}. Official runs were executed between
\textbf{14 April 2026} and \textbf{17 April 2026}.
Snapshot IDs are listed in Table~\ref{tab:models}; the three Google and two OpenAI
entries use unversioned aliases that resolve to the provider's current release at
call time.\footnote{Strict reproducibility requires pinned dated snapshots.
OpenAI publishes dated aliases (e.g.\ \texttt{gpt-4.1-2025-04-14}); Anthropic
publishes \texttt{claude-opus-4-6-<date>}. Pinned variants were not available for
all models at the time experiments were run. As a consequence, a silent provider-side
release of any aliased model between 14 and 17 April 2026 would have shifted that
model's behaviour mid-experiment in a way the run logs cannot detect; no post-hoc
API-fingerprint check (e.g.\ deterministic-prompt response stability) was
performed, and the family-by-configuration interaction observed in
Chapter~\ref{chapter:results} for the affected models (Gemini~2.5/3/3.1 and
GPT-4.1/Mini/Nano) cannot be ruled out as a snapshot-drift artefact rather
than a stable model property.}

Open-weights models are excluded from the main panel on two structural grounds: the
commercial CLM platforms surveyed in Section~\ref{sec:lit:clm} deploy proprietary APIs,
restricting external-validity claims to that deployment population; and the
long-contract tail of the corpus requires context windows of $\geq 128$K tokens,
which most open-weights models do not support. ContractEval's 19-model evaluation~\cite{liu2025contracteval} documents
a persistent proprietary-vs-open-source gap on extraction correctness; pilot work
(not formally reported) suggested this gap persists under multi-agent decomposition,
though that experiment should be understood as a scope decision rather than an empirical
claim. Open-weights extension is identified as future work in Chapter~7.

\section{Metrics}\label{sec:meth:metrics}

All metrics are formally defined in Section~\ref{sec:lit:eval}
(Equations~\ref{eq:fbeta}--\ref{eq:overextract}).

\paragraph{$F_2$ (primary).}
$\beta{=}2$ weights recall $4\times$ over precision. The asymmetry is justified by the
deployment cost structure of CLM: a missed clause passes undetected into a signed
contract, whereas a spurious extraction is caught by a reviewing lawyer. $F_2$ is the
primary hypothesis-testing metric and the basis for H1--H3.

\paragraph{$F_1$ (secondary).}
Equal-weight $F$-score, reported for comparability with prior work.
ContractEval reports $F_1$ as its headline metric~\cite{liu2025contracteval}.

\paragraph{Jaccard similarity (span quality).}
$|A \cap B| / |A \cup B|$ over token sets, computed conditional on a non-empty
prediction. Because the $F$-score TP criterion requires full span coverage,
partial extractions that cover most of a clause register as false negatives; Jaccard
gives partial credit and diagnoses near-miss behaviour independently of the
binary correctness signal.

\paragraph{Laziness rate $\lambda$ (diagnostic).}
Fraction of positive-label instances on which the model returns ``no related
clause''. The dominant failure mode of single-agent LLMs on rare CUAD categories
is silent abstention rather than wrong extraction; $\lambda$ isolates this directly.

\paragraph{Over-extraction rate $\omega$ (diagnostic).}
Fraction of negative-label instances on which the model returns a non-empty span.
Complements $\lambda$: together they decompose the false-negative and false-positive
error budgets without conflating them into a single $F$-score.

All metrics are reported pooled and per prevalence tier.

\section{Hypotheses}\label{sec:meth:hypotheses}

\paragraph{H1.} For every $C \in \{\text{M0},\, \text{M1}\}$, $C$ outperforms
$B^\star = \text{B0}$ on $F_2$ across the eleven-model panel.
Null: at least one $C$ does not outperform B0 on the panel.
Decision rule: reject the null \emph{in favour of H1} only if, \emph{for both}
$C \in \{\text{M0},\,\text{M1}\}$, a majority of the 11 per-model
$C$-vs-B0 McNemar exact contrasts is BH-significant in
favour~\cite{benjamini1995controlling} at $\alpha{=}0.05$ within the combined
$m{=}22$-test family. Per-model BH-significant positive contrasts that do
not satisfy the per-config majority condition are reported as
\emph{conditional} improvements rather than as evidence for H1 itself.
\cfgCP\ (M2) is excluded from the H1 family because it is the architecture-control
ablation for H3 (\S\ref{sec:meth:hypotheses}), not a candidate alternative
architecture; \cfgCoT\ (B1) is the chain-of-thought baseline against which the
$F_2$ comparisons are reported descriptively in Chapter~\ref{chapter:results}
but not used as the H1 reference baseline (B0 is the primary baseline per
ContractEval~\cite{liu2025contracteval}).

\paragraph{H2.} $\Delta\mathrm{Gap}(C_{\text{multi}}, B^\star) < 0$ where
$\mathrm{Gap}(C) = F_2^{\text{common}}(C) - F_2^{\text{rare}}(C)$ holds across
the eleven-model panel.
Null: H2 does not hold on the panel.
Decision rule: reject the null in favour of H2 for a given (model,
configuration) pair if the upper bound of the 95\% paired bootstrap CI on
$\Delta\mathrm{Gap}$ is strictly less than zero. H2 is supported pooled if
\emph{a majority} of the 22 (model, configuration) contrasts satisfy this rule;
otherwise H2 is rejected.

\paragraph{H3.} \cfgMA\ (M1) and \cfgCP\ (M2) are statistically distinguishable on $F_2$.
Null: $\Delta F_2(\text{M1},\,\text{M2}) = 0$.
Decision rule: reject the null (two-sided) if the McNemar exact $p$-value is below
$\alpha{=}0.05$ after BH correction ($m{=}11$, one M1 vs.\ M2 contrast per model).
The hypothesis is two-sided because direction is not pre-specified; if rejected, the
direction of the effect is reported separately because direction may be inconsistent
across the model panel.

\section{Statistical Methodology}\label{sec:meth:stats}

\paragraph{McNemar's exact test.}
For configurations $A$ and $B$ evaluated on the same instances, let
$b = |\{i : A_i \text{ correct},\, B_i \text{ wrong}\}|$ and
$c = |\{i : A_i \text{ wrong},\, B_i \text{ correct}\}|$.
The two-sided exact $p$-value is:
%
\begin{equation}\label{eq:mcnemar}
  p = 2\,\mathbb{P}\!\left[\mathrm{Bin}\!\left(b{+}c,\,\tfrac{1}{2}\right)
      \leq \min(b,c)\right]
\end{equation}
%
implemented via \texttt{scipy.stats.binomtest}~\cite{virtanen2020scipy}. The exact
form is preferred over the $\chi^2$ approximation because discordant counts are small
on rare-tier subsets~\cite{mcnemar1947note}.

\paragraph{Paired bootstrap.}
For continuous metrics ($F_2$, $F_1$, Jaccard, gap differences), Algorithm~\ref{alg:bootstrap}
gives the paired bootstrap CI~\cite{efron1979bootstrap,riezler2005pitfalls}.
We use $B{=}300$ resamples and seed the bootstrap RNG independently of the
sampler with \texttt{numpy.random.default\_rng(42)} so the data partition seed
and the bootstrap resample seed are both pinned to the same numerical value
but instantiated as separate generator objects (the bootstrap percentile CIs
in Table~\ref{tab:h1-stats} are bit-reproducible from the artefact bundle
under this convention). The percentile method's robustness to small $B$ on
bounded statistics is documented as a sensitivity threat in
Section~\ref{sec:meth:threats}.

\begin{algorithm}[h]
  \caption{Paired bootstrap confidence interval for $\Delta F_2$}
  \label{alg:bootstrap}
  \begin{algorithmic}[1]
    \Require Predictions $\hat{y}^A$, $\hat{y}^B$, labels $y$, $B{=}300$, $\alpha$
    \For{$b = 1$ \textbf{to} $B$}
      \State $\mathcal{I} \gets$ sample $n$ indices with replacement from $\{1,\dots,n\}$
      \State $\delta_b \gets F_2(\hat{y}^A_{\mathcal{I}}, y_{\mathcal{I}}) -
                           F_2(\hat{y}^B_{\mathcal{I}}, y_{\mathcal{I}})$
    \EndFor
    \State \Return $\left[\mathrm{percentile}(\delta, \alpha/2),\;
                          \mathrm{percentile}(\delta, 1{-}\alpha/2)\right]$
  \end{algorithmic}
\end{algorithm}

\paragraph{Multiple-comparison correction.}
Within each hypothesis family, Benjamini--Hochberg~\cite{benjamini1995controlling}
controls the false discovery rate (FDR) at $\alpha{=}0.05$: reject $H_0^{(i)}$ if
%
\begin{equation}\label{eq:bh}
  p_{(i)} \leq \frac{i}{m}\,\alpha,
  \quad p_{(1)} \leq \cdots \leq p_{(m)}.
\end{equation}
%
The H1 family has $m{=}22$ tests (2 multi-agent configurations
$\times$ 11 models, with B0 as reference baseline); the H3 family has
$m{=}11$ tests (one M1 vs.\ M2 contrast per model). Raw and BH-corrected $p$-values
are both reported.

\paragraph{Effect size.}
For paired binary comparisons (McNemar's), we report Cohen's
$g = b/(b{+}c) - 0.5$, the deviation of the discordant proportion from
chance~\cite{cohen1988statistical}. For continuous metrics, we report absolute
$\Delta F_2$ with the paired bootstrap 95\% CI\@. We follow Cohen's interpretive
thresholds ($g \geq 0.05$ small, $\geq 0.15$ medium, $\geq 0.25$ large;
$|\Delta F_2| \geq 0.02$ small, $\geq 0.05$ medium, $\geq 0.08$
large)~\cite{cohen1988statistical}, and report effect size alongside $p$-values in
accordance with the recommendation that the magnitude of difference is the primary
research output, not statistical significance per se~\cite{sullivan2012effect}.

\paragraph{Power.}
\textcite{card2020little} caution that NLP comparisons are frequently
under-powered. The pairwise sample size after intersection is
${\approx}\,2{,}900$--$3{,}000$ per (configuration, model) pair, of which
the discordant-pair count $b{+}c$ on M0-vs-B0 contrasts ranges from $236$
(\textit{Claude Opus~4.6}) to $1{,}030$ (\textit{Gemini~2.5~Flash}) across
the 11 models, with median $b{+}c \approx 551$. Under McNemar's exact two-sided
test at $\alpha{=}0.05$ and target power $0.80$, the median discordant-pair
count corresponds to a minimum detectable proportion of
$p_A = b/(b{+}c) \geq 0.56$ (equivalently $(b{-}c)/(b{+}c) \geq 0.12$, computed
via the binomial-distribution power formula~\cite{connor1987sample}). Mapping
discordant proportion to $F_2$ on the dominant pooled-precision regime of the
panel yields a minimum detectable $|\Delta F_2|$ of approximately $0.025$.
Effects below this threshold are treated as inconclusive rather than as
evidence of equivalence; the headline pooled spread of $0.011$ on $F_2$
across the five configurations sits within the inconclusive band by design.

\paragraph{Paired comparison protocol.}
Intersection of sample IDs is taken symmetrically when partial runs differ in coverage;
intersection size is reported with each test.

\section{Threats to Validity}\label{sec:meth:threats}

\paragraph{Internal.} Configuration parity is maintained by holding the CUAD loader,
sampling seed, prompt registry, and trace schema constant across configurations.
The M1 vs.\ M2 contrast is the most architecturally informative but not a
clean single-axis ablation: M1 differs from M2 on three jointly-varied
components: prompt-format split (one focused specialist preamble vs.\
the textual union of three), dispatch overhead (router LLM call plus
orchestrator hop), and per-call latency. Effects observed on the M1 vs.\ M2
contrast are therefore attributable to the joint architectural delta, not
to any one component in isolation; a within-M1 ablation that disables
the router decision while retaining the topology is left to future work
(\S\ref{sec:concl:outlook}).

\paragraph{External.} Single language (English), single jurisdiction (predominantly
U.S.\ commercial contracts), single corpus (CUAD). Findings on the 102-contract
test split may not extrapolate to non-CUAD clause taxonomies; the contract-length
moderation analysis (\S\ref{sec:res:length}) shows that the architectural ranking
itself shifts across length quartiles within the corpus. Open-weights
models are excluded; inference does not generalise beyond the 11 proprietary models
studied. Model snapshots are pinned; provider updates can shift behaviour.

\paragraph{Construct.} $F_2$ is a proxy for legal-review utility, not a direct
measurement. Reviewer cost of inspecting a spurious span is workflow-dependent; this
is mitigated by reporting precision, recall, $\lambda$, and $\omega$ separately so
that a reviewer with a different cost ratio can re-rank architectures.

\paragraph{Conclusion.}
The prompt-sensitivity threat is established in \S\ref{sec:lit:eval}.
We mitigate it by holding prompt templates fixed
across models per configuration, documenting them in full in Chapter~4, and treating
prompt-perturbation analysis as future work.

\paragraph{Statistical.} Bootstrap percentile CIs under-cover for small samples and
skewed statistics~\cite{efron1979bootstrap}; $p$-values are treated as decisive only
when corroborated by the McNemar exact test. The bootstrap resample count ($B{=}300$)
is below the 1{,}000 conservative recommendation~\cite{dror2018hitchhikers};
a sensitivity check on the M0-vs-B0 contrasts for the four-model Anthropic family
(\textit{Claude Haiku 4.5}, \textit{Sonnet~4}, \textit{Sonnet~4.6}, \textit{Opus~4.6};
chosen because all four runs are complete and the contrast spans the panel's
capability range from light to heavy) compared the percentile CI on
$\Delta F_2$ at $B{=}300$ versus $B{=}1{,}000$ under the same bootstrap seed
(\texttt{seed=42}, \texttt{numpy.random.default\_rng}). The maximum endpoint
shift was $0.0026$ on the lower bound and $0.0013$ on the upper bound, well
within the $\pm 0.005$ tolerance the headline reporting precision requires.
The replication script is released alongside the experiment artefacts
(\S\ref{sec:sys:opensource}). Model identity is a fixed factor; inference does
not generalise beyond the 11 models studied.
