% !TeX root = ../main.tex

\chapter{Discussion}\label{chapter:discussion}

No multi-agent configuration improves on the strongest single-agent baseline by more
than $0.011$ pooled $F_2$ points across eleven models.

\section{The Headline Finding and Its Conditions}\label{sec:disc:headline}

On the 102-contract CUAD test split under the upsampled-positive class
distribution described in \S\ref{sec:meth:sampling}, no multi-agent configuration
improves on the strongest single-agent baseline by more than $0.011$ points of
mean $F_2$ pooled across the eleven-model panel.
Under the pre-registered decision rule of Section~\ref{sec:meth:hypotheses},
H1 is rejected in its universal form: no multi-agent configuration achieves
consistent superiority across the eleven-model panel, and eleven of the 22
per-model contrasts are negative or zero. Conditional improvement exists for
individual (model, configuration) cells --- \cfgCX\ (M0) on the OpenAI family
and lighter Anthropic models --- but the pooled signal is dominated by the
model-family interaction. H2 is rejected: both pipeline decomposition
(\cfgCX, M0) and specialist routing (\cfgMA, M1) widen the rare--common $F_2$
gap relative to \cfgZS\ (B0). H3 is supported per model --- \cfgMA\ (M1) and
\cfgCP\ (M2) are statistically distinguishable on every model after BH
correction --- but the direction is inconsistent across the panel and the
pooled $F_2$ difference is zero to three decimals.

The single most informative pattern is the model-family interaction.
The Anthropic family achieves its peak under \cfgCoT\
(B1, $\overline{F_2}{=}0.851$); the Google family under \cfgZS\
(B0, $0.826$); the OpenAI family under \cfgCX\
(M0, $0.783$). The capability-tier gradient in
Table~\ref{tab:tier-gradient} is consistent with this: light- and mid-tier
models gain from M0's calibration, while heavy-tier models are better served
by B0, with the $F_2$ capability gap between Heavy and Light tiers narrowing
from $0.106$ under B0 to $0.054$ under M0 --- a $49\%$ reduction. Decomposition acts partly as a capability-equalising effect, but at a cost
(Section~\ref{sec:disc:cascade}) that is not recovered for heavy models
that already absorb long-context load in a single call. Prior work has observed
that a well-prompted single agent matches multi-agent decomposition under
stronger base models \cite{gao2025single}; the present data reproduce this:
the Anthropic Opus tier achieves its best $F_2$ under B1, not under any
decomposed architecture.

The implication for the open empirical question of when multi-agent
decomposition is justified is that, on this benchmark, the conditions are
narrower than the literature on agentic legal AI implies. A practitioner
asking whether decomposition helps a CUAD-class extraction task must answer
two prior questions: which model family, and which capability tier. On the
eleven models tested, only the OpenAI family receives a meaningful
multi-agent benefit at the pooled level, and even that benefit is dominated
in $F_2$/\$ by a B0 deployment on a cheaper Google Flash model.

\section{The Cascade as a Recall--Precision Exchange}\label{sec:disc:cascade}

Section~\ref{sec:res:cascade} identifies the classification-gate cascade as
the dominant failure mode of \cfgCX\ (M0) on the Google family:
$97.8$--$97.9\%$ of M0 false negatives on Gemini~2.5~Flash and Pro originate
at the gate rather than at the extractor. The discussion here defends the
framing of this as a recall--precision exchange rather than as pure regression.

The classifier reduces over-extraction substantially: mean $\omega$ drops
from $0.320$ under \cfgZS\ (B0) to $0.130$ under \cfgCX\ (M0), a $59\%$
reduction (Table~\ref{tab:omega}). Under $F_1$, which weights precision and
recall equally, the gate's gain is sufficient to push M0 to the highest pooled
score ($\overline{F_1}{=}0.798$ vs.\ B0's $0.737$). Under $F_2$, the gain
is insufficient to compensate the recall loss. The cascade is thus not evidence that M0 is broken, but evidence that M0 makes
a different operating-point trade. Under the pre-registered primary metric
$F_2$, the trade is unfavourable for 7 of 11 models. The practitioner question
is which operating point matches the deployment cost structure.

The cascade direction is predictable from a single-agent calibration pass.
Models with high B0 over-extraction rates (Gemini~2.5~Pro
$\omega_{\text{B0}}{=}0.456$; Gemini~2.5~Flash $0.392$) inherit a permissive
prior on negative instances; the gate amplifies the cautious prior,
suppressing recall. Models with low B0 over-extraction (GPT-4.1~Nano
$\omega_{\text{B0}}{=}0.085$) gain from the gate because their single-agent
extraction loses ground to laziness rather than to over-extraction. The
practitioner heuristic is: run B0, inspect $\omega_{\text{B0}}$, and deploy
M0 only if $\omega_{\text{B0}}$ is moderate-to-low; within the present panel,
within the present eleven-model panel, models with
$\omega_{\text{B0}} \lesssim 0.20$ gain on $F_1$ under M0, those above that
threshold do not; this boundary is indicative, not validated.

This result connects to the failure-mode taxonomy of \textcite{cemri2025why},
who identify the inter-agent verification gap as a leading cause of
multi-agent regression: the downstream extractor cannot override the upstream
classifier's silent rejection. A reflective verifier agent --- absent in
M0 --- would be the natural mitigation, and is identified as an untested
architectural pattern in Section~\ref{sec:concl:outlook}.

\section{Routing as a Deterministic Lookup}\label{sec:disc:routing}

The \cfgMA\ (M1) router agreed with a static category-to-specialist mapping on
$99.91\%$ of requests across the panel (Section~\ref{sec:res:h3}).
The natural reading is that routing in this design space reduces to a table
lookup keyed on the category name, a negative finding for the LLM-router
subspace illustrated by ChatLaw~v2~\cite{cui2024chatlaw} and
L-MARS~\cite{wang2025lmars}. Two qualifications are worth stating.

First, the result depends on three architectural commitments that do not hold
in every multi-agent design: the specialist domains were disjoint and
descriptively named; the categories carried indicators the router could read
directly from the category label; and the partition was fixed before any
model was run. In designs where domains overlap, where the category
vocabulary is open-ended, or where the routing decision must integrate
retrieved context, the router's work becomes non-trivial and routing accuracy
falls below ceiling. The generalisable claim is therefore narrower: in a
closed-vocabulary specialist-routing design, the router's contribution is
replicable by a deterministic dictionary at zero LLM cost.

Second, \cfgMA\ (M1) and \cfgCP\ (M2) differ statistically on every model
despite identical routing. The difference is not in dispatch but in prompt
format: M1 sees one focused specialist preamble per call, M2 sees the textual
union of three.
Consolidating the specialist content into a single prompt produces measurably
better span coverage (Jaccard $+0.071$ pooled, Table~\ref{tab:h3}) without
changing $F_2$. The mechanism is plausibly attentional: the union prompt
presents all three category-cluster framings simultaneously and lets the
model choose, rather than committing the dispatch decision before the
contract is read.

The combined implication for the design of legal-AI multi-agent systems is
that the value of specialist decomposition lies in the prompt content, not
in the dispatch architecture. A practitioner with a fixed specialist
taxonomy is better served by a single agent receiving the union of
specialist instructions than by a router-plus-specialists topology, and
eliminates the $23\%$ orchestrator cost premium.

\section{The $\beta{=}2$ Choice and Operating-Point Sensitivity}\label{sec:disc:metric}

The pooled architectural ranking depends on the chosen $\beta$. At
$\beta{=}1$ the ranking is \cfgCX\ $>$ \cfgMA\ $>$ \cfgZS\ $>$ \cfgCP\ $>$ \cfgCoT.
At $\beta{=}2$ the ranking is \cfgZS\ $>$ \cfgCoT\ $>$ \cfgCX\ $>$ \cfgMA\ $\approx$ \cfgCP. The architectural
conclusion of the thesis therefore depends, in part, on the $\beta$ choice.

The commitment to $\beta{=}2$ is justified in
Section~\ref{sec:meth:metrics} by the deployment cost structure of CLM
review: a missed clause carries the legal cost of an undisclosed obligation,
whereas a spurious extraction is corrected at the lower cost of attorney
review time. The empirical question of where the cost ratio actually sits is
deployment-dependent and falls outside the scope of this thesis. The
sensitivity is mitigated by reporting $F_1$, recall, precision, $\omega$,
and FN-rate alongside $F_2$, so a deployment with a different cost ratio can
re-rank architectures from the same data. The practitioner recommendation is
conditional on $\beta$: at $\beta{=}2$ deploy B0; at $\beta{=}1$ deploy M0;
below $\beta{=}1$ (where false positives are costlier than false negatives,
as in some compliance-screening workflows), M0's calibration advantage
widens further.

\section{Benchmark-to-Production Gaps}\label{sec:disc:practitioner}

The CUAD-based evaluation establishes what current architectures achieve
under idealised conditions: clean born-digital text, a fixed forty-one-category
taxonomy, expert-annotated ground truth, and a single contract per query.
\paragraph{1. Document quality is the first bottleneck.}
CUAD contracts are EDGAR filings: clean, machine-readable, consistently
formatted. Enterprise repositories contain scanned PDFs of variable OCR
quality, handwritten margin annotations, partial digitisations of legacy
contracts, and redacted sections. In the authors' assessment of production CLM deployments, document
preprocessing quality sets the primary performance ceiling --- a factor upstream
of model selection. A
model achieving high $F_2$ on CUAD text need not retain that performance on
a poorly scanned contract with OCR artefacts. This shifts the
practitioner-relevant axis of architectural design from which prompt
structure handles the contract to which preprocessing pipeline is needed
before any extraction architecture is invoked. The multi-agent design space
examined in this thesis is downstream of that question.

\paragraph{2. The task is multi-contract, not single-contract.}
CUAD evaluates a one-to-one mapping: one contract, one category. The
practitioner workflow is retrieve-then-extract: a question (``which supplier
contracts expose us to uncapped liability?'') is first resolved against a
repository of hundreds or thousands of contracts, then the relevant subset
is extracted from. This pipeline introduces three challenges that no current public benchmark
addresses: cross-contract consistency, deduplication, and result aggregation, and it shifts the architecture's value proposition from
extraction accuracy to retrieval-extraction integration. Multi-agent designs
in which a retrieval agent precedes an extraction agent are a natural fit
for this workflow, and the M0 classify-then-extract pattern is the simplest
non-trivial member of that family. The cascade failure mode of
Section~\ref{sec:disc:cascade} therefore has direct deployment relevance:
the same upstream-rejection dynamics that suppress M0 recall on Gemini~2.5
will plausibly suppress retrieval-extraction recall in production. A
controlled retrieve-then-extract benchmark on a multi-contract corpus is
identified as future work in Chapter~\ref{chapter:conclusion}.

\paragraph{3. Clause taxonomies are domain-specific and open-ended.}
The forty-one CUAD categories cover general commercial provisions.
Construction contracts contain liquidated damages tied to project milestones,
variation orders, retention terms, and defects-liability periods; SaaS
agreements contain SLA penalties, data-processing addenda, and sub-processor
notification clauses. None of these map cleanly to CUAD's category schema.
A production CLM system handles an open-ended and evolving taxonomy. The
specialist-decomposition design tested here (M1, M2) is a fixed three-domain
partition; its extensibility to an open taxonomy is the question that
matters for production, and is not what this thesis evaluates.

\paragraph{4. Production has no ground truth.}
CUAD provides expert annotations against which $F_2$, Jaccard, and $\omega$
can be computed. A production deployment has none. Evaluation proceeds
through user feedback, spot-checking, and exception triage --- none of
which benchmark metrics capture. This is the deployment-side argument for
the observability work in Chapter~\ref{chapter:system-design}: in the
absence of ground truth, the most valuable architectural property is whether
each decision can be attributed to a specific agent, prompt, and routing
branch, so that exceptions can be triaged at the source
(Section~\ref{sec:auditability}). The M0 classifier verdict and the M1
routed-specialist label are precisely the trace primitives a production
reviewer needs, even when the underlying $F_2$ is dominated by a simpler
single-agent baseline. The thesis's headline result --- that multi-agent $F_2$ does not beat
single-agent --- does not therefore translate into ``deploy single-agent in
production'', because the production utility function includes auditability,
not only $F_2$ (see Section~\ref{sec:auditability} and
Table~\ref{tab:h3} for the trace-completeness evidence).

\paragraph{5. Cross-contract reasoning is required and unmeasured.}
Production CLM questions span contracts: ``are our liability caps consistent
across all vendor agreements?''; ``which contracts are affected if
Vendor~X is acquired?''. Cross-contract reasoning requires maintaining
context across documents and detecting relationships between clauses in
different agreements. Neither CUAD nor the architectures evaluated here
address this. Hierarchical multi-agent designs --- a portfolio-level
orchestrator over per-contract extraction agents --- are a natural fit for
this workflow and are identified as future work in
Section~\ref{sec:concl:outlook}.

\paragraph{6. Integration determines deployment value.}
A system with high benchmark accuracy fails in production if it cannot
surface results inside the platform of record, in formats downstream systems
consume (risk reports, obligation trackers, renewal calendars), with
sufficient context to act upon. The post-extraction integration challenge --- transforming an extracted clause
into an actionable insight --- is largely orthogonal to architectural choice
within the extraction layer and is the axis on which
production CLM platforms compete most directly.

\medskip
\noindent
These gaps locate the CUAD evaluation rather than invalidating it: the thesis
measures one axis (extraction accuracy under fixed taxonomy and clean text)
that is necessary but not sufficient for production deployment. The
practitioner argument for multi-agent architectures therefore shifts axis:
not that decomposition raises extraction accuracy --- the data indicate it
generally does not on this benchmark --- but that decomposition provides trace primitives (structured per-agent decision
logs), modular update boundaries, and integration seams (interface boundaries
where components can be replaced or extended) that production deployments need
on axes the benchmark does not measure. This is
a defensible reason to build multi-agent CLM systems even when the benchmark
$F_2$ ranking favours single-agent baselines. It is a weaker reason than ``decomposition wins on accuracy'', and the
empirical record is better served by stating it cleanly.

\section{Threats to Validity}\label{sec:disc:threats}

The threats formalised in Section~\ref{sec:meth:threats} bound the
generalisation of these findings. Three are most consequential.

\paragraph{Single benchmark.}
The evaluation is on the 102-contract CUAD test split under the upsampled-positive
class distribution (\S\ref{sec:meth:sampling}). The findings may not extrapolate
to non-CUAD clause taxonomies, to non-English contracts, or to non-U.S.\
jurisdictions.
The contract-length moderation result
(Section~\ref{sec:res:length}, Table~\ref{tab:length-quartile}) shows that
the architectural ranking is not stable across document lengths within CUAD
itself: multi-agent helps on short contracts and hurts on long ones.
The architectural ranking observed here is therefore a ranking on this
evaluation envelope, not a universal ranking.

\paragraph{Pinned model snapshots.}
Model identifiers are pinned to specific snapshots in the run metadata.
Provider behaviour can shift under silent model updates; the family
interaction observed here may not be stable across future provider releases.
The qualitative pattern --- that architecture choice interacts with model
family more than with any within-architecture variable --- is consistent with
findings reported across task domains in the recent literature
\cite{gao2025single,cemri2025why}, which suggests the pattern is consistent across task domains even if specific
point estimates are not.

\paragraph{Closed proprietary panel.}
Open-weights models are excluded (Section~\ref{sec:meth:models}).
The proprietary--open-source gap reported by \textcite{liu2025contracteval} is
not re-examined here. The finding that multi-agent decomposition does not
improve pooled $F_2$ may not transfer to a panel of weaker open-weights
models, where decomposition's capability-equalisation effect
(Section~\ref{sec:disc:headline}) might be larger. Open-weights extension
is identified in Section~\ref{sec:concl:outlook} as the most consequential
extension of the present work.
