% !TeX root = ../main.tex

\chapter{Conclusion}\label{chapter:conclusion}

Model identity dominates architectural choice: switching configurations moves pooled $F_2$ by at most $0.011$ points, while moving across the model panel moves it by ${\approx}0.4$; the family $\times$ configuration interaction explains the residual.

\section{Summary of Contributions}\label{sec:concl:contributions}

This thesis evaluated five architectural configurations against eleven
frontier proprietary LLMs on the 102-contract CUAD test split, producing
163{,}644 paired sample-level predictions analysed under a controlled
statistical pipeline. The four contributions stated in
Chapter~\ref{chapter:introduction} are restated here with pointers to the
evidence that supports them.

\paragraph{C1: Cross-architecture benchmark.}
A cross-family cross-architecture comparison on CUAD at this scale:
five architectural configurations, eleven frontier models, 55 official runs,
analysed with paired McNemar exact tests, paired bootstrap confidence
intervals, and Benjamini--Hochberg correction. The benchmark, prompt
registry, run artefacts, and statistical pipeline are released as
open source at \url{https://github.com/ndtrung00/agentic-ai-for-clm}
(Section~\ref{sec:sys:opensource}).

\paragraph{C2: Capability-tier gradient.}
Architectural ranking is not constant across the model panel. Pooled $F_2$
across the five configurations spans a $0.011$-point band and per-model
winners are split four ways (Table~\ref{tab:h1-f2}). The capability-tier
breakdown (Figure~\ref{fig:tier-heatmap}) reveals the gradient: light- and
mid-tier models gain from \cfgCX's (M0) calibration; heavy-tier models are
dominated by \cfgZS\ (B0). No architecture is universally optimal across the eleven-model
frontier-proprietary panel evaluated here; the choice is conditional on
model family, capability tier, and the chosen $\beta$ of the $F_\beta$
metric.

\paragraph{C3: Classification-gate cascade.}
The pipeline-decomposition pattern (\cfgCX, M0) suffers a recall--precision
exchange mediated by the classification gate. On Gemini~2.5~Flash and Pro,
$97.8$--$97.9\%$ of M0 false negatives originate at the gate rather than at
the extractor (Table~\ref{tab:cascade}). The exchange's direction is
predictable from a single-agent calibration pass: models with high \cfgZS\
(B0) over-extraction lose recall to the gate; models with low B0 over-extraction
gain $F_1$ from it. Section~\ref{sec:disc:cascade} formalises the heuristic
and connects it to the inter-agent verification gap taxonomy of
\textcite{cemri2025why}.

\paragraph{C4: Routing redundancy with span-quality and grounding tradeoffs.}
The LLM router agreed with a deterministic category-to-specialist dictionary
on $99.91\%$ of requests across the panel and on $100\%$ for ten of eleven
models. \cfgCP\ (M2) matches \cfgMA\ (M1) on $F_2$ pooled to three decimals
and produces tighter spans (Jaccard $+0.071$ on ten of eleven models;
Table~\ref{tab:h3}). The architectural advantage that does survive ablation
is grounding: \cfgMA's schema-enforced specialist output achieves
$100\%$ verbatim citation of contract text on every predicted clause, against
$54\%$ for \cfgZS\ and $42\%$ for \cfgCP\ (Table~\ref{tab:grounding}). The
LLM-mediated routing call adds orchestrator cost without measurable $F_2$
benefit, but the specialist scaffold to which the router dispatches carries
the auditability surplus.

\section{Answers to the Research Questions}\label{sec:concl:rqs}

\paragraph{Main RQ.}
\emph{Can a multi-agent framework improve contract clause extraction accuracy
beyond single-agent baselines while providing superior explainability?}
On accuracy ($F_2$), multi-agent decomposition is dominated at the pooled
level by \cfgZS\ (B0), and conditionally improves only on the
OpenAI family and on lighter capability tiers. On efficiency
(cost-adjusted $F_2$), multi-agent configurations are Pareto-dominated by B0
deployments on cheaper models (Figure~\ref{fig:pareto}). On explainability,
the pipeline (\cfgCX, M0) and routed-specialist (\cfgMA, M1) topologies
provide structural auditability properties---decomposed per-agent
attribution and (under M1) schema-enforced grounding---that single-call
configurations do not produce, including chain-of-thought prompts and
reasoning-mode model variants whose rationale tokens are part of one
fused completion (Section~\ref{sec:auditability}). This is the axis on
which the practitioner argument for multi-agent architectures rests
(Section~\ref{sec:disc:practitioner}).

\paragraph{RQ1: Architectural pattern by condition.}
No pattern wins universally. Pooled $F_2$ ranks \cfgZS\ $>$ \cfgCoT\ $>$
\cfgCX\ $>$ \cfgMA\ $\approx$ \cfgCP\ within a $0.011$-point band. The
model-family interaction is the dominant signal: the Anthropic family peaks
under \cfgCoT\ (B1), the Google family under \cfgZS\ (B0), and the OpenAI
family under \cfgCX\ (M0). The prevalence-tier interaction is also large: M1
specialists lead on common-tier categories, M0 on moderate, and B0 on rare
(Figure~\ref{fig:tier-winners}). The capability-tier interaction completes the
picture: \cfgCX\ helps light and mid models, \cfgZS\ helps heavy models. The
answer to RQ1's ``under what conditions'' clause is therefore: M0 is preferable
when the model is OpenAI-family or light-tier,
$F_1$ is the primary metric, and the practitioner accepts the $1.67\times$
cost premium.

\paragraph{RQ2: Multi-agent vs.\ single-agent on accuracy and cost.}
Multi-agent configurations carry mean cost premiums of $1.23\times$
(\cfgMA, M1) to $1.67\times$ (\cfgCX, M0) over \cfgZS\ (B0) without
compensating $F_2$ gains pooled across the panel. \cfgZS\ occupies the Pareto
frontier at every cost level (Figure~\ref{fig:pareto}). The only multi-agent
cost premium that pays back in $F_2$ terms is M0 on the OpenAI family, and
even that is dominated in $F_2$/\$ by a B0 deployment on Gemini~2.5~Flash.
On $F_1$, \cfgCX\ leads pooled and the premium is partially justified for
precision-critical deployments.
This Pareto-dominance result is conditional on per-call provider billing
without prompt caching. Caching, supported by every provider in the panel,
would shrink the M0/M1 premium toward the orchestrator-hop overhead alone
and may invert the frontier on long-contract repeated-category workloads
where the system prompt and contract text are amortised across many
extraction calls (\S\ref{sec:concl:outlook}); under that cost model, RQ2 is open.

\paragraph{RQ3: Structural auditability properties.}
The parent--child call-graph schema of \cfgCX\ (M0) and \cfgMA\ (M1) provides
decomposed per-agent decision attribution (routing decision, specialist
extraction, validation outcome) and machine-checkable routing transitions
that no single-call configuration produces. The discriminating property is
structural, not the presence of rationale text: chain-of-thought prompting
(\cfgCoT, B1) and reasoning-mode model variants both emit rationale tokens,
but inside a single fused completion that lacks per-sub-decision typing
(Section~\ref{sec:auditability}). Schema conformance, measured as the
fraction of model calls for which all expected typed fields are populated,
is $100\%$ on B0/B1/M2/M0 and $95.2\%$ on M1; the residual $4.8\%$ is
attributable to provider-side truncation on long contracts
(\S\ref{sec:sys:trace-properties}). The schema is paired with enforced
grounding: \cfgMA\ reaches $100\%$ verbatim contract-text citation on every
predicted clause and \cfgCX\ reaches $89\%$, against $54\%$ for \cfgZS\ and
$42\%$ for \cfgCP\ (Table~\ref{tab:grounding}).

Three practitioner-relevant tradeoffs attend this auditability surplus.
\emph{Storage cost} scales approximately $4\times$ per M1 call (orchestrator
hop plus three specialist sub-prompts logged at inference time).
\emph{Attribution faithfulness} is bounded by the call graph: the record is
faithful to the sequence of agent calls but does not certify the model's
internal computation within any single call. \emph{Routing redundancy} adds
latency without decision diversity: the $99.91\%$ deterministic routing
result means the orchestrator layer can be replaced by a static dictionary
at zero LLM cost.

\section{Limitations}\label{sec:concl:limitations}\label{sec:disc:threats}

The limitations below bound the generalisation of these findings; they
expand the threats formalised in Section~\ref{sec:meth:threats}.

\paragraph{Single benchmark.}
The evaluation is on the 102-contract CUAD test split under the upsampled-positive
class distribution (\S\ref{sec:meth:sampling}). The findings may not extrapolate
to non-CUAD clause taxonomies, to non-English contracts, or to non-U.S.\
jurisdictions.
The contract-length moderation result
(Section~\ref{sec:res:length}, Table~\ref{tab:length-quartile}) shows that
the architectural ranking is not stable across document lengths within CUAD
itself: multi-agent helps on short contracts and hurts on long ones.
The architectural ranking observed here is therefore a ranking on this
evaluation envelope, not a universal ranking.

\paragraph{Test-set contamination.}
CUAD~\cite{hendrycks2021cuad} has been publicly hosted on HuggingFace since
2021 and is plausibly present in the pre-training data of every model in the
panel; benchmark contamination is a documented threat to the construct
validity of LLM evaluations~\cite{sainz2023nlp,magar2022data}. The present
evaluation does not separate extraction skill from memorisation, and the
$F_2$ point estimates may overstate true extraction competence. Two
observations partially constrain the threat. First, the architectural
ranking varies systematically across model families with different
documented training cut-offs (Anthropic peaks under \cfgCoT, Google under
\cfgZS, OpenAI under \cfgCX; \S\ref{sec:disc:headline}), which uniform
memorisation would not predict. Second, the M0 classification-gate cascade
(\S\ref{sec:disc:cascade}) is a recall-loss mechanism (silent upstream
rejection on positive instances) that contamination cannot generate;
memorised content would, if anything, raise recall rather than collapse it.
A held-out non-CUAD evaluation (e.g.\ a clause subset from
LegalBench or ContractNLI~\cite{koreeda2021contractnli}) is identified in
Section~\ref{sec:concl:outlook} as the highest-priority extension.

\paragraph{Pinned model snapshots.}
Model identifiers are pinned to specific snapshots in the run metadata.
Provider behaviour can shift under silent model updates; the family
interaction observed here may not be stable across future provider releases.
The qualitative pattern is consistent with findings reported across task
domains in the recent literature~\cite{gao2025single,cemri2025why}:
architecture choice interacts with model family more than with any
within-architecture variable, suggesting the pattern generalises even where
specific point estimates do not.

\paragraph{Closed proprietary panel.}
Open-weights models are excluded (Section~\ref{sec:meth:models}).
The proprietary--open-source gap reported by \textcite{liu2025contracteval} is
not re-examined here. The finding that multi-agent decomposition does not
improve pooled $F_2$ may not transfer to a panel of weaker open-weights
models, where decomposition's capability-equalisation effect
(Section~\ref{sec:disc:headline}) might be larger. Open-weights extension
is identified in Section~\ref{sec:concl:outlook} as the most consequential
extension of the present work.

\section{Future Work}\label{sec:concl:outlook}

Three axes under-determine the conclusion: untested architectural patterns,
evaluation beyond CUAD, and cost mechanisms not exercised in the experiments.

\paragraph{Untested architectural patterns.}
Three patterns absent from the five-configuration grid are expected to
behave differently. Reflective verifier agents~\cite{shinn2023reflexion}
could absorb the classification-gate cascade (\S\ref{sec:disc:cascade}) by
letting the extractor override a silent rejection on confidence-flagged
borderline cases. Tool-calling specialists that retrieve clause exemplars
or jurisdictional indicators before extracting extend the design to
open-taxonomy deployments (\S\ref{sec:disc:practitioner}). Hierarchical
orchestration (a portfolio-level orchestrator over per-contract
extraction agents) targets cross-contract reasoning, which is not exercised
by any current public benchmark.

\paragraph{Beyond CUAD.}
Four extensions widen the evaluation envelope: open-weights models, where
the proprietary--open-source gap of \textcite{liu2025contracteval} may
amplify M0's capability-equalisation effect and invert the architectural
ranking; retrieve-then-extract on a multi-contract corpus, the
production-relevant generalisation of M0; document-quality stress tests
(OCR noise, redactions, mixed-format inputs) that probe whether the
ranking survives preprocessing degradation; and formal user studies with
practising reviewers, which would close the loop on RQ3 by measuring
whether M0/M1's structural trace properties actually shorten review time.
The thesis treats these properties as a system feature, not as an outcome.

\paragraph{Cost mechanisms.}
Two cost mechanisms not exercised here would re-form the Pareto frontier
of \S\ref{sec:res:cost}. Prompt caching, supported by every provider in the
panel, would cut the marginal cost of per-category re-prompting that
dominates the B0 cost structure on long contracts and shrink the
$23$--$67\%$ multi-agent premium. Structured-output decoding would reduce
M0 classifier-call latency to where the cascade exchange becomes acceptable
on more models. The cost ranking under caching and constrained decoding
remains open.

\medskip
\noindent
On accuracy alone, the slice tested here does not justify multi-agent
deployment over a strong single-agent baseline. The case for multi-agent
architectures rests on axes the benchmark does not measure: auditability,
open-taxonomy extensibility, and cross-contract reasoning. Future work
should evaluate those axes directly.
