\chapter{\abstractname}

Contract Lifecycle Management (CLM) platforms increasingly rely on large language
models (LLMs) to automate clause extraction from commercial contracts, and the research
and vendor communities have advocated multi-agent architectures as a superior
alternative to single-prompt deployments. Despite this advocacy, no published work has
rigorously compared multi-agent architectural patterns against single-agent baselines
across multiple frontier-model families on a standardised contract-analysis benchmark
under realistic class distributions.

This thesis presents a controlled, statistically tested comparison of five
architectural configurations --- zero-shot (B1), chain-of-thought (B4), a
classify-then-extract pipeline (M0), an LLM-routed specialist system (M1), and a
consolidated specialist ablation (M6) --- evaluated across eleven frontier models from
the Anthropic, Google, and OpenAI families on 81 long contracts from the CUAD benchmark.
Evaluation uses a natural $30\%$ positive / $70\%$ negative class distribution, $F_2$
and $F_1$ as primary metrics, and paired McNemar and bootstrap statistical tests with
Benjamini--Hochberg correction.

The central finding is that no multi-agent configuration consistently outperforms the
strongest single-agent baseline on $F_2$: pooled mean $F_2$ across all eleven models
falls within a 1.1-point band ($\mathrm{B1}{=}0.801$, $\mathrm{B4}{=}0.796$,
$\mathrm{M0}{=}0.794$, $\mathrm{M1}{=}0.790$, $\mathrm{M6}{=}0.790$). Architectural
ranking varies systematically by model family: Anthropic models peak under B4
($\overline{F_2}{=}0.851$), Google models under B1 ($0.826$), and OpenAI models under
M0 ($0.783$). Two mechanistic findings explain the pooled null result. First, the
classify-then-extract pipeline (M0) suffers a \emph{classification-gate rejection
cascade} on Gemini-family models, with wrong-rejection rates of $31.7\%$--$40.9\%$
versus $1.5\%$--$2.7\%$ for the single-agent baseline, accounting for
$97.8\%$--$97.9\%$ of M0's false negatives on those models. Second, the LLM router
in M1 achieves near-perfect routing accuracy ($100\%$ on ten of eleven models),
collapsing to a trivially solvable category-name lookup; M1 and M6 are statistically
distinguishable on $F_2$ for all eleven models (BH-corrected McNemar, $p < 0.05$),
but the direction is inconsistent across families and the pooled means coincide
($\overline{F_2}{=}0.790$ for both), indicating that the routing topology itself
contributes no systematic benefit. Multi-agent configurations carry a
$1.23\times$--$1.67\times$ provider-cost premium relative to B1 with no aggregate
accuracy benefit, placing the zero-shot single-agent baseline on the Pareto frontier.

These results delineate the conditions under which multi-agent designs help, are
inert, or actively hurt in enterprise CLM, and motivate a practitioner-oriented
six-gap analysis connecting benchmark performance to production deployment reality.
